---
title: "scores simulation study"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up and utility functions for scoring

```{r}
library(scoringRules)
library(dplyr)
library(purrr)
library(ggplot2)
```

```{r}
#' Calculate interval score
#' 
#' @param y vector of observations from target distribution
#' @param l lower bound of predictive distribution
#' @param u upper bound of predictive distribution
#' @param alpha the interval [l,u] is a (1 - alpha)*100% predictive interval
#'
#' @return vector of interval scores separately for each element of x
interval_score <- function(y, l, u, alpha) {
  if(length(l) > 1 || length(u) > 1 || l > u) {
    stop("[l, u] must specify a single predictive interval.")
  }
  
  if(length(alpha) > 1 || alpha < 0.0 || alpha > 1.0) {
    stop("alpha must be a single number between 0 and 1.")
  }
  
  score = (u - l) +
    ifelse(y < l, 2.0/alpha * (l - y), 0.0) +
    ifelse(y > u, 2.0/alpha * (y - u), 0.0)
  
  return(score)
}


#' Calculate quantile score, optionally including absolute error,
#' for a parametric distribution.
#' 
#' @param y vector of observations from target distribution
#' @param alphas vector of alphas.  the score is based on several
#'   (1 - alpha)*100% predictive intervals for each alpha in alphas
#' @param ae logical.  If true, include absolute error in score
#' @param family family for distribution; e.g. 'norm' or 'nbinom'
#' @param args list of arguments for quantile function of parametric family,
#'   e.g. list(mean=0.0, sd=1.0)
#'
#' @return vector of interval scores separately for each element of x
quantile_score_parametric_dist <- function(y, alphas, ae, family, args) {
  # name of function to compute quantiles
  qfun <- paste0('q', family)
  
  # initialize score to 0
  score = rep(0.0, length(y))
  
  # interval scores for each alpha
  for(alpha in alphas) {
    args$p <- c(alpha/2, 1-alpha/2)
    interval <- do.call(qfun, args = args)
    score = score + interval_score(y, interval[1], interval[2], alpha)
  }
  
  # optionall add ae score
  if(ae) {
    args$p <- 0.5
    m <- do.call(qfun, args = args)
    score = score + abs(y - m)
  }

  return(score)
}


#' Create a list of score functions to evaluate
#'
#' @param family family for distribution; e.g. 'norm' or 'nbinom'
#' @param quantile_score_args list of lists of arguments alphas and ae for
#'   the quantile_score_parametric_dist function.  For example, 
#'   \code{list(list(alpha=c(0.05, 0.2), ae=FALSE), list(alpha=0.2, ae=TRUE))}
#'   will create two quantile score functions: the first based on a combination
#'   of interval scores from 95% and 80% predictive intervals without absolute
#'   error, and the second based on an 80% predictive interval with absolute
#'   error.
#'
#' @return named list of score functions.  Each function in the list has
#'   arguments y and family-specific arguments such as mean and sd.
assemble_score_functions <- function(family, quantile_score_args) {
  continuous_scores <- list(
    exact_crps = function(y, args) {
      args$y <- y
      do.call(
        get(paste0('crps_', family), asNamespace('scoringRules')),
        args
      )
    },
    exact_logs = function(y, args) {
      args$y <- y
      do.call(
        get(paste0('logs_', family), asNamespace('scoringRules')),
        args
      )
    }
  )
  
  quantile_scores <- purrr::map(quantile_score_args, function(qs_arg) {
    alphas <- qs_arg$alpha
    ae <- qs_arg$ae
    
    qs_name <- paste0('qs_', paste0(alphas, collapse="_"))
    if(ae) {
      qs_name <- paste0('ae_', qs_name)
    }
    
    result <- list(
      function(y, args) {
        quantile_score_parametric_dist(
          y=y,
          alphas=alphas,
          ae=ae,
          family=family,
          args=args)
      }
    )
    
    names(result) <- qs_name

    return(result)
  }) %>%
    unlist(recursive=FALSE)

  return(c(continuous_scores, quantile_scores))
}

```

# Conditions for study

```{r}
# families of target distribution: normal and negative binomial
families <- c("norm", "nbinom")

# static distribution or time-varying (dynamic); currently only static
settings <- "static"

# models to compare, specified as a list of relative location and scale
# parameters, relative to the data generating process.
# c(0.0, 1.0) means same location and scale as the data generating process.
# The location is specified as bias in units of standard deviations of the
# data generating process.
# 
# We consider:
#  * DGP
#  * wide has variance = 1.5 times variance of DGP
#  * narrow has variance = 0.5 times variance of DGP
#  * bias_wide has variance = 1.5 times variance of DGP
models <- list(
  DGP = c(0.0, 1.0),
  wide = c(0.0, sqrt(1.5)),
  narrow = c(0.0, sqrt(0.5)),
  bias_wide = c(0.2, sqrt(1.5))
)

# DGP mean and standard deviation for static setting
dgp_mean <- 100
dgp_var <- 150
```

Function to map model specifications above to family-specific arguments.

```{r}
models_to_family_args <- function(dgp_mean, dgp_var, models, family) {
  family_args <- lapply(models, function(model_adj) {
    model_mean <- dgp_mean + model_adj[1]*sqrt(dgp_var)
    model_var <- model_adj[2] * dgp_var
    if(family == 'norm') {
      return(list(
        mean = model_mean,
        sd = sqrt(model_var)
      ))
    } else if(family == 'nbinom') {
      return(list(
        mu = model_mean,
        size = (model_mean^2)/(model_var - model_mean)
      ))
    } else {
      stop('invalid family')
    }
  })
  names(family_args) <- names(models)
  
  return(family_args)
}
```

Here are tables of the parameter values for the distributions we're considering:

```{r}
norm_model_args <- models_to_family_args(
  dgp_mean, dgp_var, models, family='norm'
)
purrr::map_dfr(norm_model_args, as.data.frame) %>%
  mutate(model = names(models))
```

```{r}
nbinom_model_args <- models_to_family_args(
  dgp_mean, dgp_var, models, family='nbinom'
)
purrr::map_dfr(nbinom_model_args, as.data.frame) %>%
  mutate(model = names(models))
```

Here are some plots of the distributions we're considering:

```{r}
x_grid <- seq(from = 0, to = 200, length = 201)

to_plot <- purrr::map_dfr(
  families,
  function(family) {
    model_args <- models_to_family_args(dgp_mean, dgp_var, models, family)
    
    dfun <- paste0('d', family)
    
    return(
      purrr::map_dfr(
        seq_along(models),
        function(i) {
          data.frame(
            x = x_grid,
            density = do.call(dfun, c(list(x=x_grid), model_args[[i]])),
            family = family,
            model = names(models)[i],
            stringsAsFactors = FALSE
          )
        }
      )
    )
  }
)

ggplot(
  data=to_plot,
  mapping=aes(x=x, y=density, color=model, linetype=model)) +
  geom_line() +
  facet_wrap( ~ family, ncol=1, scales='free_y') +
  theme_bw()
```

Double checking I got the formulas right for adjustments to mean and variances by sampling.

```{r}
model_family_samples <- purrr::map_dfr(
  families,
  function(family) {
    model_args <- models_to_family_args(dgp_mean, dgp_var, models, family)
    
    rfun <- paste0('r', family)
    
    return(
      purrr::map_dfr(
        seq_along(models),
        function(i) {
          data.frame(
            x = do.call(rfun, c(list(n=10000), model_args[[i]])),
            family = family,
            model = names(models)[i],
            stringsAsFactors = FALSE
          )
        }
      )
    )
  }
)

model_family_samples %>%
  group_by(family, model) %>%
  summarize(
    sample_mean = mean(x),
    sample_var = var(x)
  )
```

# Scores to calculate

```{r}
quantile_score_args <- list(
  list(alphas = 0.05, ae = FALSE),
  list(alphas = 0.05, ae = TRUE),
  list(alphas = 0.2, ae = FALSE),
  list(alphas = 0.2, ae = TRUE),
  list(alphas = c(0.2, 0.05), ae = FALSE),
  list(alphas = c(0.2, 0.05), ae = TRUE)
)
```

# Study



```{r}
#' Perform simulation study for one combination of settings
#' Currently handles only static settings
#'
#' @param n_sim number of simulations to conduct
#' @param family family of target distribution
#' @param setting currently only 'static' supported
#' @param models list of lists with model offset parameters from dgp
#'
#' @returns data frame with scores for one combination of family and setting,
#'   all combinations of models and score functions
sim_results_one_combo <- function(
  n,
  family,
  setting = 'static',
  models) {
  if(!identical(setting, 'static')) {
    stop("only static setting implemented")
  }
  
  # scoring functions
  score_functions <- assemble_score_functions(family, quantile_score_args)
  
  # parameters for each model
  model_args <- models_to_family_args(dgp_mean, dgp_var, models, family)
  
  # generate data from dgp
  rfun <- paste0('r', family)
  y <- do.call(rfun, c(list(n=10000), model_args[[1]]))
  
  # calculate scores for all models and all scoring functions
  all_scores <- purrr::pmap_dfr(
    expand.grid(
      model_name=names(models),
      score_name=names(score_functions)
    ),
    function(model_name, score_name) {
      data.frame(
        family=family,
        setting=setting,
        i=seq_along(y),
        model_name = model_name,
        score_name = score_name,
        score = do.call(
          score_functions[[score_name]],
          args = list(y = y, args = model_args[[model_name]])
        ),
        stringsAsFactors = FALSE
      )
    }
  )
  
  return(all_scores)
}

sim_combos <- expand.grid(family=families, setting=settings, stringsAsFactors=FALSE)
sim_results <- purrr::pmap_dfr(sim_combos, sim_results_one_combo, n = 1000,models = models)
```

# Summaries

```{r}
mean_scores <- sim_results %>%
  group_by(family, setting, model_name, score_name) %>%
  summarize(mean_score = mean(score))

mean_scores %>%
  tidyr::pivot_wider(names_from = 'model_name', values_from = 'mean_score')
```

```{r}
ggplot(data = mean_scores, mapping = aes(x = score_name, y = mean_score, color = model_name, linetype = model_name, group=model_name)) +
  geom_line() +
  facet_wrap( ~ family, ncol = 1) +
  theme_bw()
```

```{r}
ggplot(data = mean_scores %>% filter(mean_score < 1000),
       mapping = aes(x = score_name, y = mean_score, color = model_name, linetype = model_name, group=model_name)) +
  geom_line() +
  facet_wrap( ~ family, ncol = 1) +
  theme_bw()
```

